\chapter{Introduction}

In the process of evolution, man strived to subjugate the surrounding world and its phenomena so that it would become the most friendly place for life and development. Over
millions of years of human development, a lot of discoveries and inventions have been made, which
have enabled man to develop the population to the point of settling the whole world. At the
moment, when the basic needs of a person living in society are met, efforts are made to optimize the
amount of work and obtaining profit using science's achievements based on new technologies.
One of the latest achievements of science is the creation of the so-called artificial intelligence and its
use in the service of man by satisfying the needs of a higher order.
Nowadays, AI is used on many levels, so a layman may overlook its impact on life and work
at the current level of development.
Society has become accustomed to automated work, quick access to information from various
areas of life, and the possibility of taking advantage of new technologies, often without thinking
about what is behind all this. Many areas of life use artificial intelligence widely at every stage of the
technological process.
Starting from education through art, culture, medicine, and exact sciences to the world of entertainment.
Having satisfied the needs of the lower order, man tries to progressively develop the possibilities of
his own creation to create structures that allow for the optimal use of acquired resources. The
development and improvement of the latest technologies make it possible to maintain the quality of
human life at a level far exceeding the standards from half a century ago.
Artificial intelligence has replaced human outwork in many areas of life, which is particularly important when performing extremely strenuous, demanding, or dangerous work. This
paper shows how artificial intelligence can be harnessed to work in a field that seems
entirely based on digressions in the area it covers. Predicting results on brokerage exchanges has so
far been based on the use of knowledge of the market as well as one's intuition. The field in which the so-called human element is indispensable will always remain beyond the
reach of automated IT systems.
In this work, attempts were made to prove that even in fields that would seem to be completely
dominated by human work, it is possible to rely on the resources of information systems.

    \par
    \bigskip
    This work aims to create a target application to learn about the possibilities of neural networks in the context of predicting the stock exchange prices. The presented project is educational and designed to enable experimentation with the use of parameters of a given neural network - such as input data size, range, epoxy quantity, et cetera. In order to allow the end user to interact with the presented application easily, the project requires creating a simple graphical interface, so the end user can easily interact with the application and quickly view the results. Experimentation and the student's interaction with the assigned field is the best way to learn. For this purpose, this work focused on simple but logical dependencies on this particular type of neural network. The aim of this project was not to present an application capable of predicting the actual stock market but rather to create an application that could improve learning and showcase the possibilities of Artificial Intelligence. The goal was not only to educate but also to inspire to act and improve the current knowledge.